TODO: A Survey of StarCraft AI Techniques

### 2012-05-11 survey.tex
review: strategy

\subsection{Strategy}
XXX Every AI needs to observe the world in order to analyze it and take decisions. In RTS games this task isn't trivial since we have to deal with imperfect information.

%\textbf{Spatial}

XXX Terrain analysis supplies the AI with chunks of abstract information about the map in order to help in making decisions. This analysis is usually performed off-line, in order to save processing time during gameplay.

Pottinger \cite{Pottinger00} described \emph{BANG!} engine implemented by Ensemble Studios for the game Age of Empires II. This engine provides terrain analysis functionalities to the game using influence maps and areas with connectivity information.

\cite{Forbus2002} showed the importance to have qualitative spatial informationfor wargames, for which they used geometric and pathfinding analysis.

\cite{Hale08} presented a 2D geometric navigation mesh generation method from expanding convex regions from seeds.

\cite{Perkins10} applied Voronoi decomposition (then pruning) %\cite{Karavelas04}
to detect regions and relevant choke points in RTS maps. This approach is implemented for StarCraft in BWTA\footnote{\url{http://code.google.com/p/bwta/}}.

%\textbf{Temporal}
% Isla 2006 Game AI Programming Wisdom 3. Charles River Media. chapter Probabilistic Target Tracking and Search Using Occupancy Maps, 379–388. ?

\cite{Hladky2008} benchmarked hidden semi-Markov models (HSMM) and particle filters in FPS units tracking. They showed that the accuracy of occupancy maps was improved using movement models (learned from the player behavior) in HSMM. 
% comparable with human experts ?

\cite{HagelbackJ08} presented a multi-agent potential fields based bot able to deal with fog of war in Tankbattle. XXX

\cite{WeberAIIDE11} used a particle model with a simple linear trajectory update to track opponent units under fog of war in StarCraft.

% {\color{ForestGreen}
% Anyone did anything here? I think only Skynet uses temporal reasoning...
% Maybe heat maps?? but I can't find any reference...
% }

%\textbf{Enemy}

XXX Opponent modeling is the process of keeping track of what the enemy is doing during a game in order to estimate the probability of it using a specific strategy.

The main methods used to these ends are case-based reasoning (CBR) and planning or plan recognition \cite{LTW,CBR_Planning,OntanonCBR,HTNPlanning,Ramirez}.\cite{LTW} used CBR to perform dynamic plan retrieval extracted from domain knowledge in Wargus (Warcraft II clone). \cite{CBR_Planning} base their real-time case-based planning (CBP) system on a plan dependency graph which is learned from human demonstration. In \cite{OntanonCBR,PlanRetrieval}, they use CBR and expert demonstrations on Wargus. They update their previous CPB approach by using ``situation assessment'' for plan retrieval. They improve the speed of CPB by using a decision tree to select relevant features. Hsieh and Sun \cite{HsiehS08} based their work on \cite{LTW} CBR model and used StarCraft replays to construct states and building sequences. Strategies are choices of building construction order in their model.

\cite{SchaddBS07} describe opponent modeling through hierarchically structured models of the opponent behavior and they applied their work to the Spring RTS (Total Annihilation clone). \cite{HTNPlanning} use hierarchical task networks (HTN) to model strategies in a first person shooter with the goal to use HTN planners. \cite{OBRecog} improve the probabilistic hostile agent task tracker (PHATT \cite{PHATT}, a simulated HMM for plan recognition) by encoding strategies as HTN. %\cite{Madeira06} 
\cite{WeberCig09} proposed ``a data mining approach to strategy prediction'' and performed supervised learning on StarCraft replays. \cite{HMMstrat_RTS_AIIDE11} used HMM to learn the transition probabilities of sequences of constructions and kept the most probables to produce probabilistic behavior models (in StarCraft). \cite{SynnaeveOpeningCig11} continued on \cite{WeberCig09} and used a Bayesian semi-supervised (labeling based on clustering) model to learn from replays and predict openings (early game strategies) from StarCraft replays, while \cite{SynnaeveAIIDE11} presents a whole game Bayesian tech-tree predictor based on unsupervised learning from replays.

Most commercial games avoid gather this information making their AI omniscient and using perfect information all the time. But sometimes they simulate some scouting tasks as Bob Fitch described in his AIIDE 2011 keynote for the game series WarCraft and StarCraft I.

{\color{ForestGreen}
Nobody had described an efficient way to scout the opponent... better opponent information better opponent modeling...
}

% \textbf{Ally}
% 
% {\color{ForestGreen}
% Anyone did anything here?
% }

\subsection{Tactics}

The most common decision making are Decision Trees. Some of the advantages of the decision trees are: simple to understand and interpret acting like a white box where we can explain why a decision is taken, and they are compatible with other decision techniques. For complex decisions we have many algorithms to generate an optimum decision tree like ID3 or C4.5, depends on the problem we will use one or another.

A good alternative are Finite State Machine. The idea behind FSM is to decompose an object's behavior into easily manageable "chunks" or states, such as "attacking an enemy", "gathering resources" or "repairing" and establish the conditions that trigger transitions between them.

>>>>

{\color{blue}
perhaps TODO ``how it's done in the industry'' regrouping the first paragraph of each of the following 3 sections?
}
%The most common decision making are Decision Trees. Some of the advantages of the decision trees are: simple to understand and interpret acting like a white box where we can explain why a decision is taken, and they are compatible with other decision techniques. For complex decisions we have many algorithms to generate an optimum decision tree like ID3 or C4.5, depends on the problem we will use one or another.

\subsection{Strategy}
%XXX Every AI needs to observe the world in order to analyze it and take decisions. In RTS games this task isn't trivial since we have to deal with imperfect information.

In the industry, strategies are often ``hard coded'' with finite state machines (FSM) \cite{FSM_AIGameProgWisdom2003}. The idea behind FSM is to decompose an object's behavior into easily manageable states, such as ``attacking'', ``gathering resources'' or ``repairing'' and establish the conditions that trigger transitions between them. Hierarchical FSM give higher (strategic) states some control over lower oned (actions). However, strategic decisions have to be made in a context, not in a vacuum. It is hard to encode dynamic, adaptive behaviors in (H)FSM due to the exponential number of transitions (between states) to account for. To this end, several approaches have been developed like hierarchical task networks, behavior trees and the comeback of STRIPS \cite{FikesSTRIPS} planning in the industry \cite{orkinGDC_FEAR}. Regardless, one cannot but notice that strategy adaptation is not the strong point of industrial RTS AI. Strategy adaptation is part of what is called opponent modeling. An adapting robot needs to keep track of what the enemy is doing in order to estimate the probability that they will use a specific strategy and adapt its own strategy to it.

In research, the main methods used for strategic decision-making are case-based reasoning (CBR) and planning or plan recognition \cite{LTW,CBR_Planning,OntanonCBR,HTNPlanning,Ramirez}. Aha et al. \cite{LTW} used CBR to perform dynamic plan retrieval extracted from domain knowledge in Wargus (Warcraft II clone). Onta\~{n}\'{o}n et al. \cite{CBR_Planning} base their real-time case-based planning (CBP) system on a plan dependency graph which is learned from human demonstration. In \cite{OntanonCBR,PlanRetrieval}, they use CBR learned from expert demonstrations on Wargus. They update their previous CBP approach by using situation assessment for retrieving the ``right'' plan, and improve the speed of CBP by using a decision tree to select relevant features. Hsieh and Sun \cite{HsiehS08} based their work on \cite{LTW} CBR model and used StarCraft replays to construct states and building sequences (``build orders''). Schadd et al. \cite{SchaddBS07} describe opponent modeling through hierarchically structured models of the opponent behavior and they applied their work to the Spring RTS (Total Annihilation clone). Hoang et al. \cite{HTNPlanning} use hierarchical task networks (HTN) to model strategies in a first person shooter with the goal to use HTN planners. Kabanza et al. \cite{OBRecog} improve the probabilistic hostile agent task tracker (PHATT \cite{PHATT}, a simulated HMM for plan recognition) by encoding strategies as HTN. %Madeira et al. \cite{Madeira06}

Weber and Mateas \cite{WeberCig09} proposed ``a data mining approach to strategy prediction'' and performed supervised learning (from buildings features) on labeled StarCraft replays. Dereszynski et al. \cite{HMMstrat_RTS_AIIDE11} used HMM to learn the transition probabilities of sequences of constructions and kept the most probables to produce probabilistic behavior models (in StarCraft). Synnaeve and Bessi\`{e}re \cite{SynnaeveOpeningCig11} extended \cite{WeberCig09} and used a Bayesian semi-supervised (labeling based on EM clustering considering appropriate features) model to learn from replays and predict openings (early game strategies) from StarCraft replays. Then \cite{SynnaeveAIIDE11}, they presented an unsupervised learning Bayesian model for tech-tree prediction, still using replays.

Most commercial games avoid gathering this information by using perfect information all the time (cheating). Sometimes they simulate some scouting tasks as Bob Fitch described in his AIIDE 2011 keynote for the WarCraft game series and StarCraft: Broodwar. \{\color{blue}XXX (perhaps move it above, to the start of subsection}

%{\color{ForestGreen}
%Nobody had described an efficient way to scout the opponent... better opponent information better opponent modeling...
%}

\subsection{Tactics}
Terrain analysis supplies the AI with chunks of abstract information about the map in order to help in making decisions. This analysis is usually performed off-line, in order to save processing time during gameplay. Pottinger \cite{Pottinger00} described the \emph{BANG} engine implemented by Ensemble Studios for the game Age of Empires II. This engine provides terrain analysis functionalities to the game using influence maps and areas with connectivity information. Forbus et al. \cite{Forbus2002} showed the importance to have qualitative spatial information for wargames, for which they used geometric and pathfinding analysis. Hale et al. \cite{Hale08} presented a 2D geometric navigation mesh generation method from expanding convex regions from seeds. Finally, Perkins \cite{Perkins10} applied Voronoi decomposition (then pruning) %\cite{Karavelas04}
to detect regions and relevant choke points in RTS maps. This approach is implemented for StarCraft in BWTA\footnote{\url{http://code.google.com/p/bwta/}}.
% Isla 2006 Game AI Programming Wisdom 3. Charles River Media. chapter Probabilistic Target Tracking and Search Using Occupancy Maps, 379–388. ?

Hladky and Bulitko \cite{Hladky2008} benchmarked hidden semi-Markov models (HSMM) and particle filters in FPS units tracking. They showed that the accuracy of occupancy maps was improved using movement models (learned from the player behavior) in HSMM. 
% comparable with human experts ?
Hagelb\"{a}ck and Johansson \cite{HagelbackJ08} presented a multi-agent potential fields based bot able to deal with fog of war in Tankbattle. XXX 
Weber et al. \cite{WeberAIIDE11} used a particle model with a simple linear trajectory update to track opponent units under fog of war in StarCraft.
